---
title: "P8105 Homework 3"
author: "Laylah Jones"
date: 2023-10-10
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(knitr)
```

# Problem 1

### Loading Data

```{r}
library(p8105.datasets)
data("instacart")
```

## Part 1. Description of instacart data

The instacart data set contains information about online grocery shopping from 2017, with `r nrow(instacart)` observations and `r ncol(instacart)` variables. Totaling at `r (nrow(instacart)*ncol(instacart))` data points. Key variables include `order_id`, `product_id`, `user_id`, if an item was `reordered`, `order_hour_of_day`, `days_since_prior_order` (ranging from `r min(pull(instacart, days_since_prior_order))` to `r max(pull(instacart, days_since_prior_order))`), `product_name`, `aisle`, and `department`.


## Part 2. Exploring the data

### How many aisles are there, and which aisles are the most items ordered from?

```{r, aisle_count}
instacart |> 
  group_by(aisle_id, aisle) |> 
  summarize(n_obs = n()) |> 
  arrange(desc(n_obs))
```

There are `r max(pull(instacart, aisle_id))` aisles. `fresh vegetables`, `fresh fruits`, and `packaged vegetable fruits` are the three most ordered items. 

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r, aisles_plot}
instacart |>  
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of Items Ordered in Each Aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r, aisle_popular}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(rank = min_rank(desc(n))) |>
  filter(rank < 4) |>
  arrange(desc(n)) |>
  knitr::kable()
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r, table_mean}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  spread(key = order_dow, value = mean_hour) |> 
  knitr::kable(digits = 2)
```

# Problem 2

### Loading Data

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

## Part 1. Data cleaning

```{r, brfss_cleaning}
brfss = brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health", 
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |> 
  mutate(
    response = fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent"))
    )
```

## Part 2. Comments on the results

### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r, states_obs}
brfss |> 
  filter(year %in% c("2002", "2010")) |> 
  group_by(year, locationabbr) |> 
  summarize(nloc = n_distinct(geo_location)) |> 
  filter(nloc >= 7) |> 
  knitr::kable()
```

Based on the results, in 2002, 6 states (CT, FL, MA, NC, NJ, PA) were observed at 7 or more locations. While in 2010, 14 states (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA) were observed at 7 or more locations.  

### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r, excellent_responses}
excell_df = 
  brfss |> 
  filter(response == "Excellent") |> 
  select(year, locationabbr, data_value) |> 
  group_by(year, locationabbr) |> 
  summarize(mean_data = mean(data_value))
```

### Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r, spaghetti_plot}
excell_df |>  
  ggplot(aes(x = year, y = mean_data, color = locationabbr)) +
  geom_line() +
  labs(title = "Average Value Over Time Within a State", x = "Years", y = "Average Data Value", color = "State Abb")
```

### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r, ny_plot}
brfss_ny = 
  brfss |> 
  select(year, data_value, response, locationdesc, locationabbr)
```


